---
title: "Reproducible Research - Week Two"
author: "merkasaur"
date: "2/19/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Reading in activity.csv file and renaming to AMD. Histogram is built. Dependencies loaded.
Steps by Day Histogram
Mean: 10766.19
Median: 10765
```{r Steps by Day Histogram, warning=FALSE}
#Dependencies
library(magrittr)
library(dplyr)
library(ggplot2)
library(readr)
#Read activity.csv.
AMD <- read.csv("activity.csv")
#Histogram
stepsdate <- AMD %>% 
      select(date, steps) %>% 
      group_by(date) %>% 
      summarise(dsteps = sum(steps)) %>% 
      na.omit()
hist(stepsdate$dsteps,xlab="Daily Steps",main="Steps by Day",col = c("gray"),breaks = 20)
#mean
##This is the mean.
mean(stepsdate$dsteps)
##This is the median.
median(stepsdate$dsteps)
```

##Time Series Plot

Time series plot is generated. Data is organized for five minute intervals and missing values are fixed. 

```{r warning = FALSE}
#Timea series plot
datatimeseries <- AMD %>%
      select(interval, steps) %>%
      na.omit() %>%
      group_by(interval) %>%
      summarize(dsteps=mean(steps))
#plot time series
ggplot(datatimeseries,aes(x=interval,y=dsteps)) +
      geom_line()
#Five minute intervals
datatimeseries[which(datatimeseries$dsteps==max(datatimeseries$dsteps)),]
#Input missing values
missingNAs <- sum(is.na(data))
#Fill in missing values
replaceNA <-function(z) replace(z, is.na(z), mean(z, na.rm=TRUE))
MD <- AMD %>%
  group_by(interval) %>%
  mutate(steps= replaceNA(steps))
```

Histogram generation of Total steps by day with Mean/Median values. 
```{r warning = FALSE}
#Histogram Total steps by day Mean/median

totalstepsbyday <- aggregate(MD$steps, by=list(MD$date), sum)
names(totalstepsbyday)[1] ="date"
names(totalstepsbyday)[2] = "completesteps"

#Mean & median of Steps by Day
summary(totalstepsbyday)
#Histogram code
hist(totalstepsbyday$completesteps, xlab="Steps", ylab = "Frequency", main = "Daily Steps Total", col= c("gray"),breaks = 20)
```
Mean Values calculated and information generated for weekday vs weekend information. GGplot actions occuring. 
```{r warning = FALSE}
# Old to New Mean
firstmean <- mean(stepsdate$dsteps, na.rm = TRUE)
secondmean <- mean(totalstepsbyday$dsteps)
#Old to new median
firstmedian <- median(stepsdate$dsteps, na.rm = TRUE)
secondmedian <- median(totalstepsbyday$dsteps)
#Weekdays vs Weekends
MD$date <- as.Date(MD$date)
MD$weekday <- weekdays(MD$date)
MD$weekend <- ifelse(MD$weekday=="Saturday" | MD$weekday=="Sunday", "Weekend", "Weekday")
MDbothwkwkd <- aggregate(MD$steps, by = list(MD$weekend, MD$interval), na.omit(mean))
names(MDbothwkwkd) <- c("weekend", "interval", "steps")
ggplot(MDbothwkwkd, aes(x=interval, y=steps))+
      geom_line()+
      facet_grid(weekend~.)+
      xlab("Interval")+
      ylab("Steps Mean")+
      ggtitle("Average steps by Interval")
```

